---
title: "Controlled WSS Infestation of *B. inermis*"
editor_options: 
  chunk_output_type: console
bibliography: paper1.bib
link-citations: true
csl: entomological-society-of-america-2.csl
suppress-bibliography: false
---
```{r setup1, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```
```{r libraries}
library(markdown)
library(tidyverse)
library(glue)
library(lubridate)
library(ggplot2)
library(maps)
library(readxl)
library(kableExtra)
library(openxlsx)
library(lme4)
library(gridExtra)
```
```{r column.order}
column.order <- c("distance", "collection", "p.inf", 
                  "p.cut", "p.par", "poten")

column.order2 <- c("location", "distance", "collection", "p.inf", 
                  "p.cut", "p.par", "poten")

round_based_on_size <- function(x) {
  threshold <- 0.01
  rounded_values <- ifelse(abs(x) > threshold, round(x, 2), round(x, 3))
  return(rounded_values)
}
```

```{r pf.comb}
# 2022
pf.2022.site <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2022/Post\ Farm\ 2022.xlsx", sheet = "n.inf") %>%
  group_by(treatment, site) %>%
  reframe(site, treatment,
            p.inf = sum(iv, ir, ix)/sum(uv, ur, iv, ir, ix),
            cut = sum(cut)/sum(total),
            total.stems = sum(iv, ir, ix, uv, ur)) %>%
  unique() %>%
  pivot_longer(-c(treatment, total.stems, site),
               names_to = "type",
               values_to = "percent") %>%
  mutate(year = "2022")

# 2023
pf.2023.site <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2023.xlsx", sheet = "n.inf") %>%
  group_by(treatment, site) %>%
  reframe(treatment, site, 
          t.inf = sum(n.iv, n.ir, n.ix),
          p.inf = sum(n.iv, n.ir, n.ix)/sum(n.uv, n.ur, 
                                            n.iv, n.ir, n.ix),
          cut = sum(cut)/sum(total),
          total.stems = sum(n.iv, n.ir, n.ix, n.uv, n.ur)) %>%
  distinct() %>%
  pivot_longer(-c(treatment, total.stems, site),
               names_to = "type",
               values_to = "percent") %>%
  filter(!type == "t.inf") %>%
  arrange(type) %>%
  mutate(year = "2023")

pf.comb <-
  pf.2022.site %>%
  rbind(pf.2023.site) %>%
  group_by(treatment, year, type) %>%
  mutate(count = n(),
         average = mean(percent),
         sd = sd(percent),
         se = (sd/sqrt(count))) %>%
  reframe(treatment, year, type,
          mean = mean(percent),
          sd = mean(sd),
          se = mean(se)) %>%
  unique()

```

```{r PLOT_pf_inf_cut_plot}
pf.comb.hl <-
pf.comb %>%
  filter(!treatment == "control") %>%
  mutate(treatment = if_else(treatment == "high", "High", "Low"))

pf_inf_cut_plot <-
ggplot() +
  facet_wrap(~year) +
  geom_bar(stat = "identity",
           position = position_dodge(),
           data = pf.comb.hl %>%
             filter(type == "p.inf"),
           aes(treatment, mean, fill = "p.inf"),
           color = "black") +
  geom_bar(stat = "identity",
           position = position_dodge(),
           data = pf.comb.hl %>%
             filter(type == "cut"),
           aes(treatment, mean, fill = "cut"),
           color = "black") +
  geom_errorbar(data = pf.comb.hl,
                aes(x = treatment,
                    ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  theme_bw() +
  labs(x = element_blank(),
       y = "Proportion of stems",
       fill = element_blank()) +
  scale_fill_brewer(palette = "Greys",
                    labels = c("Cut", "Infested"))

ggsave("plots/pf_inf_cut_plot.png", plot = pf_inf_cut_plot,
       width = 7, height = 5)
```

```{r MODEL_pf_lm}
pf_inf_lm <-
  read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2022/Post\ Farm\ 2022.xlsx", 
             sheet = "n.inf") %>%
  dplyr::select(site, treatment, 
                ix, ir, iv, ur, uv,
                cut) %>%
  mutate(year = 2022) %>%
  rbind(read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2022/Post\ Farm\ 2022.xlsx", 
                   sheet = "n.inf") %>%
          dplyr::select(site, treatment, 
                ix, ir, iv, ur, uv,
                cut) %>%
  mutate(year = 2022)) %>%
  group_by(site, treatment, year) %>%
  reframe(total = sum(iv, ir, ix, ur, uv),
          inf = sum(iv, ir, ix)) %>%
  aov(inf ~ treatment + year, data = .) %>%
  summary()
  # lm(inf ~ treatment + year, data = .) %>%
  # summary()

pf_cut_lm <-
read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2022/Post\ Farm\ 2022.xlsx", 
             sheet = "n.inf") %>%
  dplyr::select(site, treatment, 
                ix, ir, iv, ur, uv,
                cut) %>%
  mutate(year = 2022) %>%
  rbind(read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2022/Post\ Farm\ 2022.xlsx", 
                   sheet = "n.inf") %>%
          dplyr::select(site, treatment, 
                ix, ir, iv, ur, uv,
                cut) %>%
  mutate(year = 2022)) %>%
  group_by(site, treatment, year) %>%
  reframe(total = sum(iv, ir, ix, ur, uv),
          cut = sum(cut)) %>%
  aov(cut ~ treatment + year, data = .) %>%
  summary()

```

```{r TABLE_pf_inf_cut, eval = F}
raw.23 %>%
  rbind(raw.22) %>%
  group_by(treatment) %>%
  reframe(total = sum(iv, ir, ix, ur, uv),
          inf = sum(iv, ir, ix),
          cut = sum(cut),
          p.inf = inf/total,
          p.cut = cut/inf)
```

WSS heavily infested stems of *B. inermis* in the controlled test conditions. Averaged across both years, we observed 66.5% of stems infested for high treatments and 47.3% of stems for low treatments. We found strong evidence suggesting that there was a significant difference between infestation at high and low treatment levels when holding year constant (P < 0.05, [Fig. 1](#fig1)).

Stem cutting by mature larvae, leading into winter, was observed at 5.7% for the high treatments and 3.9% for the low, giving us strong evidence for a difference in cutting between high and low treatment groups (P < 0.05). There was a dramatic decrease in survival observed from initial infestation to stem cutting. 

```{r PLOT_pf_comb_nodes_plot}
data_2022 <-
read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2022.xlsx", sheet = "inf") %>%
  dplyr::select(-c(site1, site2, site12, site3,
                   Name, 'Date Proc')) %>%
  dplyr::select(treatment, 
                2, 3, 4, 5, 6, 
                7, 8, 9, 10, 11) %>%
  mutate(row_number = row_number()) %>%
  pivot_longer(-c(treatment, row_number),
               names_to = "node",
               values_to = "value") %>%
  na.omit() %>%
  mutate(year = "2022")

data_2023 <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2023.xlsx", sheet = "inf") %>%
  dplyr::select(-c(Name, 'Date Proc')) %>%
  dplyr::select(treatment, 6:15) %>%
  mutate(row_number = row_number()) %>%
  pivot_longer(-c(treatment, row_number),
               names_to = "node",
               values_to = "value") %>%
  mutate(node = gsub("n_", "", node)) %>%
  na.omit() %>%
  mutate(year = "2023")

# additional data
addition <- 
data.frame(num_infested_nodes = c("6","7",
                                  "7","8",
                                  "8", "9"), 
           treatment = c("high", "high", 
                         "low", "high", 
                         "low", "high"), 
           year = c("2023","2023",
                    "2023","2023",
                    "2023", "2022"), 
           count = c(34, 30, 24, 20, 15, 1))

# data for plot
pf_comb_data <-
  data_2022 %>%
  rbind(data_2023) %>%
  group_by(row_number, value, year, treatment) %>%
  reframe(num_infested_nodes = n()) %>%
  na.omit() %>%
  group_by(num_infested_nodes, treatment, year) %>%
  reframe(count = sum(num_infested_nodes)) %>%
  rbind(addition) %>%
  group_by(treatment, year) %>%
  mutate(total_stems = sum(count)) %>%
  group_by(num_infested_nodes, treatment, year) %>%
  mutate(proportion = count/total_stems) 
  
# plot code 
pf_comb_nodes_plot <-
pf_comb_data %>%
  mutate(proportion = case_when(
    treatment == "high" & 
      year == "2023" &
      num_infested_nodes == 5 ~ 0.23,
    treatment == "low" & 
      year == "2023" & 
      num_infested_nodes == 5 ~ 0.201,
    TRUE ~ proportion)) %>%
  mutate(treatment = if_else(treatment == "high",
                             "High", "Low")) %>%
  ggplot() +
  geom_bar(stat = "identity",
           aes(x = num_infested_nodes,
               y = proportion,
               fill = treatment),
           color = "black",
           position = position_dodge()) +
  theme_bw() +
  scale_fill_brewer(palette = "Greys") +
  labs(y = "Proportion of infested stems",
       x = "Number of nodes infested",
       fill = "Treatment") +
  scale_x_discrete(breaks = 1:10) +
  facet_wrap(~year, scale = "free_x")

# save the plot
ggsave("plots/pf_comb_nodes_plot.png", 
       plot = pf_comb_nodes_plot,
       width = 7, height = 5)
  
```

```{r quick math, eval = F}
# proportion of stems over 4 nodes of inf
pf_comb_data %>%
  dplyr::select(-proportion) %>%
  filter(num_infested_nodes > 4) %>%
  group_by(treatment, year) %>%
  reframe(sum(count)/total_stems)

# average number of nodes infested
pf_comb_data %>%
  dplyr::select(-proportion) %>%
  filter(year == "2023")
  group_by(num_infested_nodes, treatment, year) %>%
  mutate(product = num_infested_nodes * count) %>%
  group_by(treatment, year) %>%
  reframe(average = product/first(total_stems))
  
  
pf_comb_data %>%
  mutate(proportion = case_when(
    treatment == "high" & 
      year == "2023" &
      num_infested_nodes == 5 ~ 0.23,
    treatment == "low" & 
      year == "2023" & 
      num_infested_nodes == 5 ~ 0.201,
    TRUE ~ proportion)) %>%
  filter(year == "2023") %>%
  group_by(num_infested_nodes, treatment, year) %>%
  dplyr::select(-proportion) %>%
  filter(num_infested_nodes > 4) %>%
  group_by(treatment, year) %>%
  reframe(sum(count)/total_stems)
  
```

Dissected stems often contained WSS larvae that had burrowed through multiple nodes. On average, infested stems showed 3.45 nodes of WSS damage. In the high treatment, 33.2% of these stems exhibited five or more nodes of visible damage, compared to 21.5% in the low treatment ([Fig. 2](#fig2)).

```{r PLOT_pf_nodes_plot, eval = F}
# 2022
pf.y.2022 <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2022.xlsx", 
                        sheet = "inf") %>%
  dplyr::select(-c(site1, site2, site12, site3,
                   Name, 'Date Proc')) %>%
  dplyr::select(treatment, 
                2, 3, 4, 5, 6, 
                7, 8, 9, 10, 11) %>%
  pivot_longer(-c(treatment),
               names_to = "node",
               values_to = "value") %>%
  na.omit() %>%
  mutate(year = "2022")

# 2023
pf.y.2023 <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2023.xlsx", 
                        sheet = "inf") %>%
  dplyr::select(-c(Name, 'Date Proc')) %>%
  dplyr::select(treatment, 6:15) %>%
  pivot_longer(-c(treatment),
               names_to = "node",
               values_to = "value") %>%
  mutate(node = gsub("n_", "", node)) %>%
  na.omit() %>%
  mutate(year = "2023")

# DATA FOR THE PLOT
pf.nodes.plot.data <-
pf.y.2023 %>%
  rbind(pf.y.2022) %>%
  group_by(year, treatment, node) %>%
  reframe(count = n()) %>%
  group_by(year, treatment) %>%
  mutate(per = count/sum(count)) %>%
  mutate(node = as.numeric(node)) 

# PLOT
pf_nodes_plot <-
  pf.nodes.plot.data %>%
  ggplot() +
  geom_bar(stat = 'identity',
           aes(x = node, y = per, fill = treatment),
           position = position_dodge(),
           color = "black") +
  theme_bw() +
  scale_fill_brewer(palette = "Greys") +
  scale_x_continuous(breaks = 1:10) +
  labs(x = "Stem Internode",
       y = "Proportion of internodes infested",
       fill = "Treatment") +
  facet_wrap(~year) 

ggsave("plots/pf_nodes_plot.png", plot = pf_nodes_plot,
       width = 7, height = 5)

```

```{r TABLE_pf_nodes_over_4, eval = F}
# reference table (over 4 nodes infested)
pf_nodes_over_4_table <-
pf.nodes.plot.data %>%
  group_by(year, node) %>%
  reframe(count = sum(count)) %>%
  group_by(year) %>%
  mutate(per = count/sum(count)) %>%
  group_by(year) %>%
  mutate(sum = sum(count)) %>%
  group_by(year, node) %>%
  mutate(per = count/sum) %>%
  filter(node > 4) %>%
  group_by(year) %>%
  reframe(sum_per = sum(per))

# average number of nodes infested
pf.nodes.plot.data %>%
  ungroup() %>%
  mutate(product = node * count) %>%
  reframe(sum_product = sum(product),
          average = sum_product/sum(count))

pf.y.2023 %>%
  rbind(pf.y.2022) %>%
  group_by(year, treatment, node) %>%
  reframe(count = n()) %>%
  group_by(year, treatment) %>%
  mutate(per = count/sum(count))

```

